---
title: "US_Airline_Network_Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### load packages
```{r load packages}
library("dplyr")
library("ggplot2")
library("igraph")
```

### Load data
```{r}
load("airline16.Rdata")
load("airline1516.Rdata")
load("airline_sample.Rdata")
```

### Nodes Attributes

#### Calculate daily volume of airports as an attribute of nodes
```{r}
airport_out = airline1516 %>% select(ORIGIN,FL_DATE) %>% group_by(ORIGIN,FL_DATE) %>% summarise(frequency = n()) %>% rename(airport = ORIGIN) %>% group_by(airport) %>% summarize(volume = mean(frequency))
airport_in = airline1516 %>% select(DEST,FL_DATE) %>% group_by(DEST,FL_DATE) %>% summarise(frequency = n()) %>% rename(airport = DEST) %>% group_by(airport) %>% summarise(volume = mean(frequency))
airport_volume = rbind(airport_in,airport_out) 
airport_volume = airport_volume%>% group_by(airport) %>% summarize(volume_in_out = sum(volume))
airport_volume$airport = as.character(airport_volume$airport)
airport_volume = airport_volume %>% arrange(airport)
```


#### Identify city of airports as an attribute of nodes
```{r}
airport_o_city = airline1516 %>% select(ORIGIN,ORIGIN_CITY_NAME) %>% distinct(ORIGIN,ORIGIN_CITY_NAME) %>% rename(airport = ORIGIN,city_name = ORIGIN_CITY_NAME)
airport_d_city = airline1516 %>% select(DEST,DEST_CITY_NAME) %>% distinct(DEST,DEST_CITY_NAME) %>% rename(airport = DEST,city_name = DEST_CITY_NAME)
airport_city = rbind(airport_o_city,airport_d_city)
airport_city = airport_city %>% distinct(airport,city_name)
airport_city$airport = as.character(airport_city$airport)
airport_city$city_name = as.character(airport_city$city_name)
airport_city = airport_city %>% arrange(airport)
```

#### Identify state of airports as an attribute of nodes
```{r}
airport_o_state = airline1516 %>% select(ORIGIN,ORIGIN_STATE_NM) %>% distinct(ORIGIN,ORIGIN_STATE_NM) %>% rename(airport = ORIGIN,state_name = ORIGIN_STATE_NM)
airport_d_state = airline1516 %>% select(DEST,DEST_STATE_NM) %>% distinct(DEST,DEST_STATE_NM) %>% rename(airport = DEST,state_name = DEST_STATE_NM)
airport_state = rbind(airport_o_state,airport_d_state)
airport_state = airport_state %>% distinct(airport,state_name) 
airport_state$airport = as.character(airport_state$airport)
airport_state$state_name = as.character(airport_state$state_name)
airport_state = airport_state %>% arrange(airport)
```

#### Airport Delay Rate
```{r}
Dep_Delay_Rate = airline1516 %>% mutate(have_depdelay =ifelse(DEP_DELAY_NEW>0,1,0) ) %>% select(ORIGIN,have_depdelay)
Dep_Delay_Rate = Dep_Delay_Rate[complete.cases(Dep_Delay_Rate),]
Dep_Delay_Rate = Dep_Delay_Rate %>% rename(airport = ORIGIN) %>% arrange(airport) %>% group_by(airport) %>% summarize(frequency = sum(have_depdelay)/n())
Dep_Delay_Rate$airport = as.character(Dep_Delay_Rate$airport)
Dep_Delay_Rate = Dep_Delay_Rate %>% arrange(airport)
```

#### Airport Delay Average time histogram
```{r}
Dep_top10 = airline1516 %>% select(ORIGIN, DEP_DELAY_NEW) %>% filter(DEP_DELAY_NEW>10) %>% rename(airport = ORIGIN) %>% group_by(airport) %>% summarise(avg_delay_time = mean(DEP_DELAY_NEW)) %>% top_n(10,avg_delay_time)
Delay_top10_airport = Dep_top10[1]
Delay_top10_airport_hist = airline1516 %>% select(ORIGIN,DEP_DELAY_NEW) %>% filter(ORIGIN %in% Delay_top10_airport$airport, DEP_DELAY_NEW>0)
ggplot(Delay_top10_airport_hist,aes(x = ORIGIN, y = DEP_DELAY_NEW))+
  geom_boxplot()
```


#### Airport Delay Average time(Regard delay under 10 minutes as on time)
```{r}
Dep_Delay_Time = airline1516 %>% select(ORIGIN, DEP_DELAY_NEW) %>% filter(DEP_DELAY_NEW>0) %>% rename(airport = ORIGIN) %>% group_by(airport) %>% summarise(avg_delay_time = mean(DEP_DELAY_NEW))
Dep_Delay_Time$airport = as.character(Dep_Delay_Time$airport)
Dep_Delay_Time = Dep_Delay_Time %>% arrange(airport)
```
#### geographic information
```{r}
geo_airport = read.csv("geo_airport.csv")
```

#### airport carrier delay
```{r}
airport_carrier_delay = airline1516 %>% select(ORIGIN,FL_DATE,CARRIER,DEP_DELAY_NEW) %>% mutate(delay = ifelse(DEP_DELAY_NEW>0,1,0))
airport_carrier_delay$delay[which(is.na(airport_carrier_delay$delay))] = 0
airport_carrier_delay = airport_carrier_delay %>% group_by(ORIGIN,CARRIER) %>% summarize(rate = sum(delay)/n())
airport_carrier_delay_max = airport_carrier_delay %>% group_by(ORIGIN) %>% mutate(high = max(rate))
airport_carrier_delay_max = airport_carrier_delay_max %>% mutate(is_high = ifelse(rate == high, 1,NA)) %>% na.omit %>% select(ORIGIN,CARRIER,rate) %>% rename(high_carrier = CARRIER,high_rate = rate)
airport_carrier_delay_max$ORIGIN = as.character(airport_carrier_delay_max$ORIGIN)
airport_carrier_delay_max$high_carrier = as.character(airport_carrier_delay_max$high_carrier)
airport_carrier_delay_max = arrange(airport_carrier_delay_max,ORIGIN)

airport_carrier_delay_min = airport_carrier_delay %>% group_by(ORIGIN) %>% mutate(low = min(rate))
airport_carrier_delay_min = airport_carrier_delay_min %>% mutate(is_low = ifelse(rate == low, 1,NA)) %>% na.omit %>% select(ORIGIN,CARRIER,rate) %>% rename(low_carrier = CARRIER,low_rate = rate)
airport_carrier_delay_min$ORIGIN = as.character(airport_carrier_delay_min$ORIGIN)
airport_carrier_delay_min$low_carrier = as.character(airport_carrier_delay_min$low_carrier)
airport_carrier_delay_min = arrange(airport_carrier_delay_min,ORIGIN)

airport_carrier_delay = full_join(airport_carrier_delay_max,airport_carrier_delay_min,by = "ORIGIN")
```







### Edge Attributes

#### Distance
```{r}
Mean_Distance = airline1516 %>% select(ORIGIN,DEST,DISTANCE) %>% group_by(ORIGIN,DEST) %>% summarise(mean_distance = mean(DISTANCE))

Mean_Distance$ORIGIN = as.character(Mean_Distance$ORIGIN)
Mean_Distance$DEST = as.character(Mean_Distance$DEST)
Mean_Distance = Mean_Distance %>% arrange(ORIGIN,DEST)
```

#### Airline daily  average frequency
```{r}
Airline_Frequency = airline1516 %>% select(ORIGIN,DEST,FL_DATE) %>% group_by(ORIGIN,DEST,FL_DATE) %>% summarize(frequency = n()) %>% group_by(ORIGIN,DEST) %>% summarise(daily_frequency = mean(frequency))
Airline_Frequency$ORIGIN  = as.character(Airline_Frequency$ORIGIN)
Airline_Frequency$DEST  = as.character(Airline_Frequency$DEST)
Airline_Frequency = Airline_Frequency %>% arrange(ORIGIN,DEST)
```

#### Airline delay rate
```{r}
Airline_delay_rate = airline1516 %>% mutate(have_depdelay =ifelse(DEP_DELAY_NEW>0,1,0) ) %>% select(ORIGIN,DEST,have_depdelay)
Airline_delay_rate$have_depdelay[which(is.na(Airline_delay_rate$have_depdelay))] = 0
Airline_delay_rate = Airline_delay_rate %>%  group_by(ORIGIN,DEST) %>% summarize(rate = sum(have_depdelay)/n())
Airline_delay_rate$ORIGIN = as.character(Airline_delay_rate$ORIGIN)
Airline_delay_rate$DEST = as.character(Airline_delay_rate$DEST)
Airline_delay_rate = Airline_delay_rate %>% arrange(ORIGIN,DEST)
```

#### Airline delay average time
```{r}
Airline_delay_average_time = airline1516  %>% select(ORIGIN,DEST,DEP_DELAY_NEW) 
Airline_delay_average_time = Airline_delay_average_time %>% filter(DEP_DELAY_NEW>0) %>%  group_by(ORIGIN,DEST) %>% summarize(delay_time = mean(DEP_DELAY_NEW))
Airline_delay_average_time$ORIGIN = as.character(Airline_delay_average_time$ORIGIN)
Airline_delay_average_time$DEST = as.character(Airline_delay_average_time$DEST)
Airline_delay_average_time = Airline_delay_average_time %>% arrange(ORIGIN,DEST)
airline_order = Airline_Frequency[,1:2]
airline_order$delay_time = 0

Airline_delay_average_time = rbind(Airline_delay_average_time,airline_order)
Airline_delay_average_time = Airline_delay_average_time %>% group_by(ORIGIN,DEST) %>% summarise(delay = sum(delay_time))
```

#### Airline carrier delay rate
```{r}
airline_carrier_delay = airline1516 %>% select(ORIGIN,DEST,FL_DATE,CARRIER,DEP_DELAY_NEW) %>% mutate(delay = ifelse(DEP_DELAY_NEW>0,1,0))
airline_carrier_delay$delay[which(is.na(airline_carrier_delay$delay))] = 0
airline_carrier_delay = airline_carrier_delay %>% group_by(ORIGIN,DEST,CARRIER) %>% summarize(rate = sum(delay)/n())
airline_carrier_delay_max = airline_carrier_delay %>% group_by(ORIGIN,DEST) %>% mutate(high = max(rate))
airline_carrier_delay_max = airline_carrier_delay_max %>% mutate(is_high = ifelse(rate == high, 1,NA)) %>% na.omit %>% select(ORIGIN,DEST,CARRIER,rate) %>% rename(high_carrier = CARRIER,high_rate = rate) %>% group_by(ORIGIN,DEST) %>% summarise(high_rate = mean(high_rate),high_carrier = first(high_carrier))
airline_carrier_delay_max$ORIGIN = as.character(airline_carrier_delay_max$ORIGIN)
airline_carrier_delay_max$DEST = as.character(airline_carrier_delay_max$DEST)
airline_carrier_delay_max$high_carrier = as.character(airline_carrier_delay_max$high_carrier)
airline_carrier_delay_max = arrange(airline_carrier_delay_max,ORIGIN,DEST)

airline_carrier_delay_min = airline_carrier_delay %>% group_by(ORIGIN,DEST) %>% mutate(low = min(rate))
airline_carrier_delay_min = airline_carrier_delay_min %>% mutate(is_low = ifelse(rate == low, 1,NA)) %>% na.omit %>% select(ORIGIN,DEST,CARRIER,rate) %>% rename(low_carrier = CARRIER,low_rate = rate) %>% group_by(ORIGIN,DEST) %>% summarise(low_rate = mean(low_rate),low_carrier = first(low_carrier))
airline_carrier_delay_min$ORIGIN = as.character(airline_carrier_delay_min$ORIGIN)
airline_carrier_delay_min$DEST = as.character(airline_carrier_delay_min$DEST)
airline_carrier_delay_min$low_carrier = as.character(airline_carrier_delay_min$low_carrier)
airline_carrier_delay_min = arrange(airline_carrier_delay_min,ORIGIN,DEST)

airline_carrier_delay = full_join(airline_carrier_delay_max,airline_carrier_delay_min,by = c("ORIGIN","DEST"))
```

#### airline carrier delay average time
```{r}
airline_carrier_delay_time_max = airline1516 %>% select(ORIGIN,DEST,CARRIER,DEP_DELAY_NEW) %>% filter(DEP_DELAY_NEW>0) %>% group_by(ORIGIN,DEST,CARRIER) %>% summarize(avg_time = mean(DEP_DELAY_NEW)) %>% group_by(ORIGIN,DEST) %>% mutate(high_time = max(avg_time)) %>% mutate(is_high = ifelse(avg_time == high_time,1,NA)) %>% na.omit() %>% select(ORIGIN,DEST,CARRIER,high_time) %>% rename(high_carrier = CARRIER) %>% group_by(ORIGIN,DEST) %>% summarise(high_time = mean(high_time),high_carrier = first(high_carrier))
airline_carrier_delay_time = airline_carrier_delay[,1:2]
airline_carrier_delay_time = left_join(airline_carrier_delay_time,airline_carrier_delay_time_max,by = c("ORIGIN","DEST"))


airline_carrier_delay_time_min = airline1516 %>% select(ORIGIN,DEST,CARRIER,DEP_DELAY_NEW) %>% filter(DEP_DELAY_NEW>0) %>% group_by(ORIGIN,DEST,CARRIER) %>% summarize(avg_time = mean(DEP_DELAY_NEW)) %>% group_by(ORIGIN,DEST) %>% mutate(low_time = min(avg_time)) %>% mutate(is_low = ifelse(avg_time == low_time,1,NA)) %>% na.omit() %>% select(ORIGIN,DEST,CARRIER,low_time) %>% rename(low_carrier = CARRIER) %>% group_by(ORIGIN,DEST) %>% summarise(low_time = mean(low_time),low_carrier = first(low_carrier))
airline_carrier_delay_time =left_join(airline_carrier_delay_time,airline_carrier_delay_time_min,by = c("ORIGIN","DEST"))
```


### Core Network
```{r}
airline = Airline_Frequency[,1:2]
airline_network = graph_from_data_frame(airline,directed = TRUE)
```

### Insert attributes into core network
```{r}
V(airline_network)$volume = airport_volume$volume_in_out
V(airline_network)$city = airport_city$city_name
V(airline_network)$state = airport_state$state_name
V(airline_network)$delay_rate = Dep_Delay_Rate$frequency
V(airline_network)$avg_delay_time = Dep_Delay_Time$avg_delay_time
V(airline_network)$latitude = geo_airport$latitude
V(airline_network)$longitude = geo_airport$longitude
V(airline_network)$lat = geo_airport$latitude
V(airline_network)$lng = geo_airport$longitude
V(airline_network)$delay_carrier_high = airport_carrier_delay$high_carrier
V(airline_network)$delay_carrier_high_rate = airport_carrier_delay$high_rate
V(airline_network)$delay_carrier_low = airport_carrier_delay$low_carrier
V(airline_network)$delay_carrier_low_rate = airport_carrier_delay$low_rate
E(airline_network)$distance = Mean_Distance$mean_distance
E(airline_network)$daily_frequency = Airline_Frequency$daily_frequency
E(airline_network)$deley_rate = Airline_delay_rate$rate
E(airline_network)$airline_avg_delay_time = Airline_delay_average_time$delay
E(airline_network)$linedelay_ca_high = airline_carrier_delay$high_carrier
E(airline_network)$linedelay_ca_high_rate = airline_carrier_delay$high_rate
E(airline_network)$linedelay_ca_low = airline_carrier_delay$low_carrier
E(airline_network)$linedelay_ca_low_rate = airline_carrier_delay$low_rate

E(airline_network)$linedelaytime_ca_high = airline_carrier_delay_time$high_carrier
E(airline_network)$linedelaytime_ca_high_time = airline_carrier_delay_time$high_time
E(airline_network)$linedelaytime_ca_low = airline_carrier_delay_time$low_carrier
E(airline_network)$linedelaytime_ca_low_time = airline_carrier_delay_time$low_time
```

### Save as GML file
```{r}
write_graph(airline_network,"airline_network.gml","gml")
airline_network = read_graph("airline_network.gml","gml")

```
```{r}
write.csv(airport_volume,"list.csv")

```

### Hub analysis
```{r}
t = table(degree(airline_network))
plot(t)

```
### delay analysis
```{r}

```

